### 2015-10-25 paper/lrec/lrec2016_listenmaa_claessen.tex
more cleanup

%
% File nodalida2015.tex
%
% Contact beata.megyesi@lingfil.uu.se
%
% Based on the instruction file for EACL 2014
% which in turn was based on the instruction files for previous 
% ACL and EACL conferences.

\documentclass[11pt]{article}
\usepackage{nodalida2015}
\usepackage{times}
\usepackage{mathptmx}
\usepackage{fixltx2e}
\usepackage{url}
\usepackage{multirow}
\usepackage{latexsym}
\usepackage{cite}
\usepackage{graphicx}
%\usepackage{slantsc}
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{authordate1-4}
\usepackage{multirow}
%\special{papersize=210mm,297mm} % to avoid having to use "-t a4" with dvips 
%\setlength\titlebox{6.5cm}  % You can expand the title box if you really have to

\usepackage{fontspec}
\usepackage{wasysym}

\newcommand{\mainfontfamily}{Palatino}
\setmainfont[
  Ligatures=TeX,
  Scale=0.9,
]{\mainfontfamily}

\newfontfamily\scshape[Letters=SmallCaps, Numbers=Uppercase]{Palatino
  Small Caps}

\newcommand{\monofontfamily}{Monaco} % Monaco
\setmonofont[Scale=0.8]{\monofontfamily}

\usepackage{color}
\newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}

\title{Analysing Constraint Grammars with a SAT-solver}

\author{Inari Listenmaa \and Koen Claessen \\
 Chalmers University of Technology, Gothenburg, Sweden \\
 {\tt \{inari,koen\}@chalmers.se} }

\date{\today}

\begin{document}
\maketitle

\begin{abstract}
%\small
We describe a method for
 analysing Constraint Grammars by encoding the
rules in SAT.
Our tools can detect internal conflicts or redundancies in a grammar,
as well as generate examples to demonstrate the effect of some rule or
rule set.
This can help users to diagnose and improve their grammars.
No corpus is required, only a morphological lexicon.
%We test our tools with grammars from 100-1000 rules.
\end{abstract}

% Introduction & problem description
\input{problem_desc}

% Implementation details; very brief for the abstract
\input{implementation}

% Preliminary evaluation and future work
\input{eval_futurework}

\section*{Acknowledgments}
We thank Eckhard Bick for the idea to apply SAT to CG analysis.

\bibliographystyle{acl}
\bibliography{cg}

\end{document}