### 2016-02-11 paper/lrec/lrec2016_listenmaa_claessen.tex
use LREC official stylesheet

\documentclass[10pt, a4paper]{article}
\usepackage{lrec2016}
\usepackage{multibib}
\newcites{languageresource}{Language Resources}
\usepackage{graphicx} % for eps graphics
\usepackage{epstopdf}
\usepackage[latin1]{inputenc}

% Inari's additions
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{color}
\newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
\usepackage{units} % for nicefrac
\usepackage{wasysym} % for clock

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Analysing Constraint Grammars with a SAT-solver}

\name{Inari Listenmaa, Koen Claessen}

\address{University of Gothenburg, Chalmers University of Technology \\
         Gothenburg, Sweden \\
         inari@chalmers.se, koen@chalmers.se\\}

\abstract{
We describe a method for analysing Constraint Grammars by encoding the
rules in SAT.
Our tools can detect internal conflicts or redundancies in a grammar,
as well as generate examples to demonstrate the effect of some rule or
rule set.
This can help users to diagnose and improve their grammars.
No corpus is required, only a morphological lexicon. \\ 
\newline 
\Keywords{Constrain Grammar, SAT, Grammar Analysis} }

\begin{document}

\maketitleabstract

% Introduction & problem description
\input{problem_desc}

% Implementation details; very brief for the abstract
\input{implementation}

% Preliminary evaluation and future work
\input{eval_futurework}

\section*{Acknowledgments}
We thank Eckhard Bick for the idea to apply SAT to CG analysis, and Francis Tyers for insightful discussions.

\section{Bibliographical References}
\label{main:ref}

\bibliographystyle{lrec2016}
\bibliography{cg}

% \section{Language Resource References}
% \label{lr:ref}
% \bibliographystylelanguageresource{lrec2016}
% \bibliographylanguageresource{cg}

\end{document}