### 2016-02-29 paper/lrec/lrec2016_listenmaa_claessen.tex
small changes to abstract

\documentclass[10pt, a4paper]{article}
\usepackage{lrec2016}
\usepackage{multibib}
\newcites{languageresource}{Language Resources}
\usepackage{graphicx} % for eps graphics
\usepackage{epstopdf}
\usepackage[latin1]{inputenc}

% Inari's additions
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{color}
\newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
\usepackage{units} % for nicefrac
\usepackage{wasysym} % for clock

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Analysing Constraint Grammars with a SAT-solver}

\name{Inari Listenmaa, Koen Claessen}

\address{University of Gothenburg, Chalmers University of Technology \\
         Gothenburg, Sweden \\
         inari@chalmers.se, koen@chalmers.se\\}

\abstract{
We describe a method for analysing and improving Constraint Grammars (CG).
% by encoding the rules in SAT.
Traditional CG engines execute the rules in an imperative and
sequential manner, which makes it impossible to track interaction
between rules.
In contrast, we represent CG rules as logic constraints, which
can then be tested for contradictions and inconsistencies.
Our tool can detect internal conflicts or redundancies in a grammar,
as well as generate examples to demonstrate the effect of some rule or
rule set.
This can help users to diagnose and improve their grammars.
No corpus is required, only a morphological lexicon. \\ 
\newline 
\Keywords{Constrain Grammar, SAT, Grammar Analysis} }

\begin{document}

\maketitleabstract

% Introduction & problem description
\input{problem_desc}

% Implementation details; very brief for the abstract
\input{implementation}

% Preliminary evaluation and future work
\input{eval_futurework}

\section*{Acknowledgments}
We thank Eckhard Bick for the idea to apply SAT to CG analysis, and Francis Tyers for insightful discussions.

\section{Bibliographical References}
\label{main:ref}

\bibliographystyle{lrec2016}
\bibliography{../cg}

% \section{Language Resource References}
% \label{lr:ref}
% \bibliographystylelanguageresource{lrec2016}
% \bibliographylanguageresource{cg}

\end{document}