### 2016-03-03 paper/lrec/lrec2016_listenmaa_claessen.tex
abstract

\documentclass[10pt, a4paper]{article}
\usepackage{lrec2016}
\usepackage{multibib}
\newcites{languageresource}{Language Resources}
\usepackage{graphicx} % for eps graphics
\usepackage{epstopdf}
\usepackage[latin1]{inputenc}

% Inari's additions
\usepackage[colorlinks=true,citecolor=blue,urlcolor=blue]{hyperref}
\usepackage{color}
\newcommand{\todo}[1]{{\color{cyan}\textbf{[TODO: }#1\textbf{]}}}
\usepackage{units} % for nicefrac
\usepackage{wasysym} % for clock

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Analysing Constraint Grammars with a SAT-solver}

\name{Inari Listenmaa, Koen Claessen}

\address{University of Gothenburg, Chalmers University of Technology \\
         Gothenburg, Sweden \\
         inari@chalmers.se, koen@chalmers.se\\}

\abstract{
We describe a method for analysing Constraint Grammars (CG) that can detect internal conflicts and redundancies in a given grammar, without the need for a corpus. The aim is for grammar writers to be able to automatically diagnose, and then manually improve their grammars. Our method works by automatically translating the given grammar into logical constraints that are analysed by a SAT-solver. We have evaluated our analysis on a number of non-trivial grammars and found inconsistencies.
\newline 
\Keywords{Constraint Grammar, SAT, Grammar Analysis} }

\begin{document}

\maketitleabstract

% Introduction & problem description
\input{problem_desc}

% Implementation details; very brief for the abstract
\input{implementation}

% Preliminary evaluation and future work
\input{eval_futurework}

\section*{Acknowledgments}
We thank Eckhard Bick for the idea to apply SAT to CG analysis, and Francis Tyers for insightful discussions.

\section{Bibliographical References}
\label{main:ref}

\bibliographystyle{lrec2016}
\bibliography{../cg}

% \section{Language Resource References}
% \label{lr:ref}
% \bibliographystylelanguageresource{lrec2016}
% \bibliographylanguageresource{cg}

\end{document}